He-Man: A domain-specific language for lightweight concurrency

### 2012-03-08 paper/paper.tex
Rewrite the event loop model section.

Networking applications require a very constrained form of
concurrency. Because I/O (both network and disk) latency dominates the
computation time, each client's thread of execution spends most of its
time waiting on I/O. In particular, we would like to always yield to a
client no longer blocked on I/O; we need not yield to blocked clients,
nor do we (usually) need arbitrary interleaving of execution.

%% XXX: do we really want to talk about this?
As of version 2.5.44, Linux has supported \t{epoll} for non-blocking
multiplexing I/O. A server can register any number of file descriptors with
\t{epoll\_ctl}, and a call to \t{epoll\_wait} returns the ready descriptors.

In the event loop model, all blocking I/O must occur in a single loop around
\t{epoll\_wait}. Each thread is stored as a \t{struct} containing thread-local
state, including the next instruction pointer, and each registered descriptor
has a pointer to the thread waiting on it. Instead of performing an I/O call, a
thread simply registers a new descriptor and returns back to the main loop;
whenever a descriptor is ready, the main loop jumps to the instruction pointer
of the thread waiting on it. Spawning a new thread requires only allocating a
new thread \t{struct}, and registering it in the main loop.

Each thread must be split into one function for each jump target, and
thread-local state must be accessed via the thread \t{struct} shared between
these functions. Furthermore, because I/O only occurs in the main loop, these
functions must return immediately after registering new descriptors. As a
result, each thread is split up into quite a few functions in this paradigm.
Because this translation from threaded to event-based is quite mechanical, we
are able to perform it automatically in the He-Man back-end.

>>>>

Network applications require a very constrained form of
concurrency. Because in most network applications, I/O (both network
and disk) latency dominates the computation time, each client's thread
of execution spends most of its time waiting on I/O. In particular, we
would like to always yield to a client no longer blocked on I/O; we
need not yield to blocked clients, nor do we (usually) need arbitrary
interleaving of execution.

In the event loop model, network I/O is done using non-blocking
\t{read} and \t{write} calls and all of the blocking for I/O is
centralized in a single call to a multiplexing system call such as
\t{select} or \t{epoll}. The multiplexing call will block until there
is some I/O ready to be performed and indicate to the application
which file descriptors are ready. From this, the application then
needs to decide how handle these events. This requires, for each
connection, having some sort of state machine to track what state the
connection is in and what needs to happen next, as well as other local
state.

One common scheme for doing this involves registering a function
pointer with the event loop to be called on an event. Typically each
connection is also associated with a structure of connection local
data. Instead of making a blocking system call, then, an application
registers an event with the event loop along with the function to call
when it occurs, and then returns back to the loop.

The advantage of this scheme is that it is highly scalable and can be
efficient: handling a new connection typically just involves
allocating a new structure and registering it with the event loop. The
disadvantage is that now network facing code needs to be scattered
through many small functions, with all access to connection state
mediated through a shared structure.

The event loop model is then, essentially, built around closure
converted trampolining continuation passing style. These are well
studied problems and can easily be done mechanically in the He-Man
backend.
